---
title: "Analysis of information source and health protective behaviours"
author: "Huiyi Hu"
date: '2022-04-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
library(knitr)
library(psych)
library(likert)

edu_levels <- c(
  "Bachelors or equivalent", 
  "Masters or other postgraduate qualification", 
  "Doctorate", 
  "Other") 


use_levels <- c(
  "None", 
  "Not very much", 
  "A fair amount", 
  "A lot") 


trust_levels <- c(
  "Very distrustful", 
  "Distrustful", 
  "Neutral", 
  "Trustful", 
  "Very trustful") 


likert_levels <- c(
  "Strongly disagree", 
  "Disagree", 
  "Neutral", 
  "Agree", 
  "Strongly agree")    

hpb_levels <- c(
  "Not at all", 
  "Sometimes", 
  "Often", 
  "Always")


    
d_infohpb <- read.csv('../data/info-hpb-database.csv',colClasses = "factor")


d_infohpb$age <- factor(d_infohpb$age,ordered = TRUE)
d_infohpb$gender <- factor(d_infohpb$gender,ordered = TRUE)
d_infohpb$ukstu <- factor(d_infohpb$ukstu)
d_infohpb$edqual <- factor(d_infohpb$edqual,
                             levels=edu_levels,
                             ordered = TRUE)
d_infohpb$vcamp <- factor(d_infohpb$vcamp)
d_infohpb$infomost <- factor(d_infohpb$infomost)
d_infohpb$use_news <- factor(d_infohpb$use_news,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_gov <- factor(d_infohpb$use_gov,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_uni <- factor(d_infohpb$use_uni,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_smed <- factor(d_infohpb$use_smed,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_web <- factor(d_infohpb$use_web,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_ff <- factor(d_infohpb$use_ff,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$use_doc <- factor(d_infohpb$use_doc,
                             levels=use_levels,
                             ordered = TRUE)
d_infohpb$tru_news <- factor(d_infohpb$tru_news,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_gov <- factor(d_infohpb$tru_gov,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_uni <- factor(d_infohpb$tru_uni,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_smed <- factor(d_infohpb$tru_smed,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_web <- factor(d_infohpb$tru_web,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_ff <- factor(d_infohpb$tru_ff,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$tru_doc <- factor(d_infohpb$tru_doc,
                             levels=trust_levels,
                             ordered = TRUE)
d_infohpb$covid_aff <- factor(d_infohpb$covid_aff,
                             levels=likert_levels,
                             ordered = TRUE)
d_infohpb$info_worrry <- factor(d_infohpb$info_worrry,
                             levels=likert_levels,
                             ordered = TRUE)
d_infohpb$covid_sick <- factor(d_infohpb$covid_sick,
                             levels=likert_levels,
                             ordered = TRUE)
d_infohpb$hp_hand <- factor(d_infohpb$hp_hand,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_supp <- factor(d_infohpb$hp_supp,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_clean <- factor(d_infohpb$hp_clean,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_mask <- factor(d_infohpb$hp_mask,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_win <- factor(d_infohpb$hp_win,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_sdist <- factor(d_infohpb$hp_sdist,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_dietexc <- factor(d_infohpb$hp_dietexc,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_avdpt <- factor(d_infohpb$hp_avdpt,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_onlime <- factor(d_infohpb$hp_onlime,
                             levels=hpb_levels,
                             ordered = TRUE)
d_infohpb$hp_home <- factor(d_infohpb$hp_home,
                             levels=hpb_levels,
                             ordered = TRUE)


d_infohpb <- d_infohpb %>%
  mutate(
    across(
      .cols = c(starts_with('use'),starts_with('tru'),starts_with('hp')),
      .fns = as.numeric,
      .names = '{.col}_coded'
    )
   )


names(d_infohpb)[48:57]
d_infohpb$hpb_scores <- rowSums(d_infohpb[ , c(48:57)])


dt_infouse <- d_infohpb[, c("use_news","use_gov","use_uni","use_smed","use_web","use_ff","use_doc")]


dt_infotrust <- d_infohpb[, c("tru_news","tru_gov","tru_uni","tru_smed","tru_web","tru_ff","tru_doc")]

dt_hpb <- d_infohpb[, c("hp_hand","hp_supp","hp_clean","hp_mask","hp_win","hp_sdist","hp_dietexc","hp_avdpt","hp_onlime","hp_home")]

```

# 1. Demographic characteristics of the sample
```{R}
summary(d_infohpb)

```

## 1.1 Age
```{r}
d_infohpb %>%
  count(age)

d_infohpb %>%
  ggplot(aes(x = age)) +
  geom_bar()+
  ggtitle('Gender of respondents')
```
## 1.2 Gender
```{r}
d_infohpb %>%
  count(gender)

d_infohpb %>%
  ggplot(aes(x = gender)) +
  geom_bar()
```
## 1.3 Current educational qualifications
```{r}
d_infohpb %>%
  count(edqual)

d_infohpb %>%
  ggplot(aes(x = edqual)) +
  geom_bar() +
  coord_flip()
```
## 1.4 Visiting campus
```{r}
d_infohpb %>%
  count(vcamp)

d_infohpb %>%
  ggplot(aes(x = vcamp)) +
  geom_bar() +
  coord_flip()
```
## 1.5 Health protective behaviour
```{r}
coded_summary<- d_infohpb %>%
  select(ends_with('_coded'))


# Mean value---Health protective behaviours
  coded_summary %>%
  select(starts_with('hp_'))%>%
  lapply(
    mean, na.rm = TRUE
  ) 

# Median---Health protective behaviours
coded_summary %>%
  select(starts_with('hp_'))%>%
  lapply(
    median, na.rm = TRUE
  )

# Standard deviation ---Health protective behaviours
coded_summary %>%
  select(starts_with('hp_'))%>%
  lapply(
    SD, na.rm = TRUE
  )

```

# 2. Descriptive Statistics
## 2.1(a)Information Source used most
```{r}
d_infohpb %>%
  count(infomost)


d_infohpb %>%
  ggplot(aes(x =  infomost)) +
  geom_bar() +
  ggtitle('Number of Covid-19 information source mostly used')+
  coord_flip()
```

## 2.2 hp behaviour scores (grouped by information source used most)
```{R}
ggplot(data = d_infohpb, mapping = aes(x = infomost, y = hpb_scores)) + 
  geom_boxplot(add = "jitter") +
  scale_x_discrete(name = NULL, limits = c(
    "News Media    (e.g. news from TV, radio, newspaper and/or through their websites)","Government Sources & Other Official Sources   (e.g. GOV.UK, NHS, WHO)","Communications from Universities   (e.g. emails, posters around campus, messages from tutors)","Social Media   (e.g. Facebook, Twitter, YouTube, Weibo, Xiaohongshu)","Websites   (e.g. web surfing on non-official sources, podcasts)","Family, Friends or Collegues","Doctor or Medical Provider"),
                   labels= c("News Media","Government & other official sources","Communications from Universities","Social Media","Websites of non-official source","Family, Friends or Collegues","Doctor or Medical Provider"))+
  coord_flip()
```


## 2.3 covidworry~hpb_scores(by infomost)
```{R}

d_infohpb <- d_infohpb %>%
  mutate(
    across(
      .cols = c(covid_aff,info_worrry,covid_sick),
      .fns = as.numeric,
      .names = '{.col}_coded'
    )
   )

names(d_infohpb)[59:61]
d_infohpb$worry_level <- rowSums(d_infohpb[ , c(59:61)])


ggplot(d_infohpb,aes(hpb_scores, worry_level))+
  geom_point(aes(color = infomost)) +
  scale_x_continuous(labels = NULL) +
  geom_smooth(se = FALSE) +
  scale_y_continuous(labels = NULL)+
  theme(legend.position = "bottom")


```



## 2.4 covid-19 info received from Information Sources
```{R}
tibble(
  d_infohpb$use_news,
  as.numeric(d_infohpb$use_news)) %>%
  kable

d_infohpb$use_news_coded %>% as.numeric %>% median(na.rm = TRUE)


likert(dt_infouse) %>%
  plot(type='bar')
```

## 2.5 Trust in information source
```{R}

likert(dt_infotrust) %>%
  plot(type='bar')+
  guides(color = guide_legend(override.aes = list(size = 5),keyheight =1,title = NULL))+
  scale_color_manual(breaks = 5:1,
                     labels = rev(c("Very Distrustful","Distrustful", "Neutral","Trustful","Very Trustful")), 
                     values = c("tan3","tan1","grey85","steelblue1","steelblue3"))


```

## 2.6 Health Protective Behaviours
```{R}
likert(dt_hpb) %>%
  plot(type='bar')


# visualize dependent variables
ggplot(df, aes(x = hpb_scores, y = ..density..)) + geom_histogram(fill = "blue", colour = "grey60", size = 0.2, alpha = 0.2) + geom_density()



```


## 2.6.1 geom_point: hpb
```{r}
# Create new dataset with mean value of hpb
coded_hpb<- 
  coded_summary %>%
  select(starts_with('hp_'))%>%
  lapply(
    mean, na.rm = TRUE
  ) 

coded_hpb <- tibble(
  Item = coded_hpb %>% names,
  mean = coded_hpb %>% unlist
)

coded_hpb

coded_hpb %>%
  ggplot(
    aes(x = Item, y = mean)
  ) +
  geom_point() +
  scale_x_discrete(name = NULL, limits = c(
    "hp_hand_coded","hp_supp_coded","hp_clean_coded","hp_mask_coded","hp_win_coded","hp_sdist_coded","hp_dietexc_coded","hp_avdpt_coded","hp_onlime_coded","hp_home_coded"),
                   labels= c("Wash hand with sanitizer","Take dietary supplements","Disinfect surface","Wear face masks","Stay in well ventilated areas","Keep social distance","Maintain healthy diet and exercise","Reduce use of public transport","Arrange meeting people online","Stay home when feeling unwell"))+
  coord_flip()




```

# 3. Statistical Testing
## 3.1 Test of demographic variables
```{R}


#change gender into dummy variable
d_infohpb <- d_infohpb %>%
  mutate(
    female = gender == 'Female')

# t test gender
d_infohpb %>%
  t.test(
    hpb_scores~gender,
    var.equal = FALSE,
    data = .
  )
```
```{r}

#person cor test
#cor test-gender(Female)

d_infohpb %>%
  mutate(female=female %>% as.numeric) %>%
  cor.test(~ female+hpb_scores, data= .)

#cor test-gender(Age)
d_infohpb %>%
  mutate(age=age %>% as.numeric) %>%
  cor.test(~ age+hpb_scores, data= .)

#cor test-gender(Eduqual)
d_infohpb %>%
  mutate(edqual=edqual %>% as.numeric) %>%
  cor.test(~ edqual+hpb_scores, data= .)

```


## 3.2 Welch's Anova test of info-source variables
```{r}

 
#change [social media use] into dummy variable
d_infohpb <- d_infohpb %>%
  mutate(
    social_media = infomost == 'Social Media   (e.g. Facebook, Twitter, YouTube, Weibo, Xiaohongshu)')


# t test social media
d_infohpb %>%
  t.test(
    hpb_scores~social_media,
    var.equal = FALSE,
    data = .
  )



```

## Welch's ANOVA
```{R}

oneway.test(hpb_scores ~ social_media, data = d_infohpb, var.equal = FALSE)

```

 
```{R}
wilcox.test(use_news_coded ~ gender, data=d_infohpb,paired=FALSE) 
wilcox.test(use_gov_coded ~ gender, data=d_infohpb,paired=FALSE)

wilcox.test(hpb_scores ~ gender, data=d_infohpb,paired=FALSE)

```
## H1: Test of hpb_scores ~ demographic variables
```{R}
wilcox.test(hpb_scores ~ gender, data=d_infohpb,paired=FALSE)
kruskal.test(hpb_scores ~ age, data = d_infohpb) 
kruskal.test(hpb_scores ~ edqual, data = d_infohpb) 

 
```


#4. Correlation
## 4.1 Correlation Matrix-Spearsman
```{R}
df<- d_infohpb %>%
  select(age,female,edqual,
          use_news,use_gov,use_uni,use_smed,use_web,use_ff,use_doc,
              tru_news,tru_gov,tru_uni,tru_smed,tru_web,tru_ff,tru_doc,
         hpb_scores)


library("Hmisc")
df[] <- lapply(df, as.numeric)

             
res2 <- rcorr(as.matrix(df[]),type='spearman')
res2_p <- res2[["P"]]
res2_m <- res2[["r"]]

df %>%
  ggplot(
    aes(x = Item, y = mean)
  ) +
  geom_point() +
  coord_flip()



M<-cor(df)
head(round(M,2))
corrplot(M, method="circle",tl.col="black")
corrplot(M, method="color",tl.col="black")



# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(df)
head(p.mat[, 1:5])

# Leave blank on no significant coefficient
corrplot(M, type="upper", order="hclust", 
         p.mat = p.mat, sig.level = 0.5, insig = "blank")

## Final Version of correlation matrix
library(RColorBrewer)
corrplot(M,tl.col="black",
         col=brewer.pal(n=10, name="RdBu"))

m %>%
  ggplot(
    aes(x = Item, y = mean)
  ) +
  geom_point() +

```

# 5. Linear Regression Model
```{R}
#Model 1
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,edqual,infomost,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
  lm(hpb_scores~age+female+edqual+infomost,
     data= .) %>%
  summary


#Model 2
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,edqual,infomost,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
  lm(hpb_scores~age+female+edqual+infomost
                +
                use_news+use_gov+use_uni+use_smed+
                use_web+use_ff+use_doc,
     data= .) %>%
  summary


#Model 3
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,edqual,infomost,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
  lm(hpb_scores~age+female+edqual+infomost+
                tru_news+tru_gov+tru_uni+tru_smed+
                tru_web+tru_ff+tru_doc,
     data= .) %>%
  summary


#Model 4
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,edqual,infomost,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
  lm(hpb_scores~age+female+edqual+infomost+
                +
                use_news+use_gov+use_uni+use_smed+
                use_web+use_ff+use_doc
                +
                tru_news+tru_gov+tru_uni+tru_smed+
                tru_web+tru_ff+tru_doc,
     data= .) %>%
  summary


  ## Model test-1 only use
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,infomost,covid_aff,covid_sick,info_worrry,
                social_media,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
      lm(hpb_scores~age+female
                +
                use_news+use_gov+use_smed+use_ff
                ,
     data= .) %>%
  summary
  
 ## Model test
d_infohpb %>%
  mutate(
    across(
      .cols = c(age,female,edqual,infomost,covid_aff,covid_sick,info_worrry,
                social_media,
                starts_with('use'),
                starts_with('tru'),
                starts_with('hp')),
      .fns = as.numeric)) %>%
      lm(hpb_scores~age+female+infomost+edqual
              +use_ff+tru_smed
                ,
     data= .) %>%
  summary

```


# rank-order model
```{R}

##use and trust level were dichotomised by coding the top2 levels as 1, the others as 0
    
df_logic <- read.csv('../data/df_logic.csv')


##fit-use
fit.use <- lm(hpb~female+age+edqual+
                use_news+use_gov+use_uni
                +use_smed+use_web+use_ff+use_doc
                , data = df_logic)
summary(fit.use) 

```

```{r}
##fit-trust
fit.trust <- lm(hpb~female+age+edqual
                 +tru_news+tru_gov+tru_uni
                +tru_smed+tru_web+tru_ff+tru_doc, data = df_logic) 
summary(fit.trust)
```

```{R}


fit.model1 <- lm(hpb~female+age
                 +tru_uni
                +use_smed+tru_ff, data = df_logic) 
summary(fit.model1)
```

```{r}
##fit-all
fit.full <- lm(hpb~female+age+edqual
                +use_news+use_gov+use_uni
                +use_smed+use_web+use_ff+use_doc
                +tru_news+tru_gov+tru_uni
                +tru_smed+tru_web+tru_ff+tru_doc, data = df_logic)
summary(fit.full)
```

```{r}

df_logic2 <- read.csv('../data/df_logic2.csv')
##fit-all
```

```{r}
fit.full2 <- lm(hpb~female+age+edqual
                +use_news+use_gov+use_uni
                +use_smed+use_web+use_ff+use_doc
                +tru_news+tru_gov+tru_uni
                +tru_smed+tru_web+tru_ff+tru_doc, data = df_logic2)
summary(fit.full2)
```